Application of Equal Local Levels to Improve Q-Q Plot Testing Bands with \proglang{R} Package \pkg{qqconf}

### 2021-06-08 article/manuscript.Rnw

A number of methods have been created towards this end. Naively, one could put pointwise testing bands onto Q-Q plots, which conduct $\alpha$ level tests on each order statistic of the sample. This method is clearly insufficient, as it ignores the multiple testing problem and thus provides no guarantee on the Type I error rate of testing the global null hypothesis that all of the data come from the specified null distribution. To appropriately deal with this multiple testing problem, the Komolgorov-Smirnov (KS) statistic is commonly used to create joint testing bands \citep{kolmogoroff1941confidence, smirnov1944approximate}. While this method ensures the correct Type I error by using the distribution of the KS statistic to create joint testing bands, this test suffers from very low power under a variety of reasonable alternatives since it has low sensitivity to deviation in the tails of the null distribution \citep{aldor2013power, berk1979goodness}. A more promising approach would be to apply the equal local levels method (ELL) to create testing bands. This method, originally introduced by Berk and Jones and later developed by Gontscharuk et al. as an alternative testing procedure to higher-criticism and KS, conducts a test with Type I error rate $\eta$ on each order statistic of the sample such that the Type I error rate of the global test is some desired level $\alpha$ \citep{berk1979goodness, gontscharuk2015intermediates, gontscharuk2016goodness, gontscharuk2017asymptotics}. Because this test conducts an $\eta$ level test on each order statistic, it can be used to create testing bands that are agnostic to any alternative distribution. While the ELL method has never previously been specifically applied to creating testing bands on Q-Q plots, we suggest it is a sensible choice for a standard test in this use case, as it does not bias the user towards deviations of the sample in any particular part of the null distribution.

@@

Currently, there is no available software that allows a user to quickly create plausible and robust testing bands for Q-Q plots on real data. While the base-\proglang{R} package \pkg{stats} provides functionality for creating a Q-Q plot to compare a sample against the normal distribution, it doesn't give the user the ability to create Q-Q plots for other distributions or to put testing bands on those plots. The package \pkg{qqplotr} provides a number of helpful additions to the base-\proglang{R} functionality, including the ability to create Q-Q plots for a variety of reference distributions and a robust method for testing bands on the normal distribution \citep{qqplotr}. Specifically, for the normal distribution \pkg{qqplotr} implements a simulation based method called tail sensitive (TS) testing bands, originally developed by Alder-Noiman \citep{aldor2013power}.  However, this package does not provide adequate testing bands for non-normal reference distributions and is not able to efficiently create testing bands for large sample sizes because it uses simulation.

@@

In what follows, we introduce the \proglang{R} package \pkg{qqconf}. With the aid of a novel recursive calculation developed below, \pkg{qqconf} can quickly provide ELL testing bands for comparing even very large samples to any reference distribution with a CDF implemented. In addition to these testing bands, \pkg{qqconf} provides a variety of plotting functionalities that allow the user to easily visualize where any deviation of the sample from the null distribution may occur. In section \hyperref[sec:methods]{2}, we introduce the methods required for the efficient recursive calculation mentioned above. In section \hyperref[sec:methods]{3}, we discuss the behavior of ELL bands in large sample sizes and an approximation to these bands that makes them very easy to calculate for even very large samples ($n > 5 \cdot 10^{6}$). In section \hyperref[sec:examples]{4}, we introduce \pkg{qqconf} and demonstrate its functionality on a few multiple testing scenarios including including Type I error calibration and genetic studies. Finally, we conclude in section \hyperref[sec:discussion]{5} with a discussion of limitations the of ELL and \pkg{qqconf} as well as further directions for creating testing bands on Q-Q plots.

@@

This is known as ``equal local levels.''

@@

@@

>>>>

A number of methods have been created towards this end. Naively, one could put pointwise testing bands onto Q-Q plots, which conduct an $\alpha$ level test on each order statistic of the sample. This method is clearly insufficient, as it ignores the multiple testing problem and thus provides no guarantee on the Type I error rate of testing the global null hypothesis that all of the data come from the specified null distribution. To appropriately deal with this multiple testing problem, the Komolgorov-Smirnov (KS) statistic is commonly used to create joint testing bands \citep{kolmogoroff1941confidence, smirnov1944approximate}. While this method ensures the correct Type I error by using the distribution of the KS statistic to create joint testing bands, this test suffers from very low power under a variety of reasonable alternatives since it has low sensitivity to deviation in the tails of the null distribution \citep{aldor2013power, berk1979goodness}. A more promising approach would be to apply the equal local levels method (ELL) to create testing bands. This global testing method, originally introduced by Berk and Jones and later developed by Gontscharuk et al. as an alternative testing procedure to higher-criticism and KS, conducts a test with Type I error rate $\eta$ on each order statistic of the sample such that the Type I error rate of the global test is some desired level $\alpha$ \citep{berk1979goodness, gontscharuk2015intermediates, gontscharuk2016goodness, gontscharuk2017asymptotics}. Previously, Alder-Noiman applied this method to create testing bands for Q-Q plots for the normal distribution using simulation \citep{aldor2013power}. Because this test conducts an $\eta$ level test on each order statistic and thus does not bias the user towards deviations of the sample from any particular part of the null distribution, we suggest it is a sensible choice for a standard method in creating testing bands for Q-Q plots.

@@

Currently, there is no available software that allows a user to quickly create plausible and robust testing bands for Q-Q plots on real data. While the base-\proglang{R} package \pkg{stats} provides functionality for creating a Q-Q plot to compare a sample against the normal distribution, it doesn't give the user the ability to create Q-Q plots for other distributions or to put testing bands on those plots. The package \pkg{qqplotr} provides a number of helpful additions to the base-\proglang{R} functionality, including the ability to create Q-Q plots for a variety of reference distributions and a robust method for testing bands on the normal distribution \citep{qqplotr}. Specifically, for the normal distribution \pkg{qqplotr} implements Alder-Noiman's simulation based method (which is referred to in the package as tail sensitive (TS) testing bands). However, this method does not provide testing bands for non-normal reference distributions and is not able to efficiently create testing bands for large sample sizes because it uses simulation.

@@

In what follows, we introduce the \proglang{R} package \pkg{qqconf}. With the aid of a novel recursive calculation developed below, \pkg{qqconf} can quickly provide ELL testing bands for comparing even very large samples to any reference distribution with a CDF implemented. In addition to these testing bands, \pkg{qqconf} provides a variety of plotting functionalities that allow the user to easily visualize where any deviation of the sample from the null distribution may occur. In section \hyperref[sec:methods]{2}, we introduce the methods required for the efficient recursive calculation mentioned above. In section \hyperref[sec:methods]{3}, we discuss the behavior of ELL bands in large sample sizes and an approximation to these bands that makes them very easy to calculate for even very large samples ($n > 5 \cdot 10^{6}$). In section \hyperref[sec:examples]{4}, we introduce \pkg{qqconf} and demonstrate its functionality on a few multiple testing scenarios including including Type I error calibration and genetic studies. Finally, we conclude in section \hyperref[sec:discussion]{5} with a discussion of some limitations of the ELL method and \pkg{qqconf} as well as further directions for creating testing bands on Q-Q plots.

@@

We refer to this as the ELL method, as mentioned in the Introduction.

@@

\subsection{Testing for Normality of Regression Residuals}
When performing an ordinary least squares regression (OLS), it is common to assume that the error terms are idependentally drawn from a normal distribution. Specifically, suppose that we have $n$ observations $\{\mathbf{x_{i}}, y_{i}\}_{i=1}^{n}$, where each $y_{i}$ is a scalar and each $\mathbf{x_{i}}$ is a $1xp$ vector of regressors. Then, we model each $y_{i}$ as
\begin{equation*}
  y_{i} = \mathbf{x_{i}}^{T}\boldsymbol\beta + \epsilon_{i}
\end{equation*}
where $\mathbf{\boldsymbol\beta}$ is a $1xp$ vector of regression parameters to be estimated and we assume
\begin{equation}
  \epsilon_{1}, ..., \epsilon_{n} \distras{iid} N(0, \sigma^{2}).
\end{equation}
Once $\boldsymbol\beta$ and $\sigma^{2}$ are estimated (typically via maximum liklihood), for each $y_{i}$ we define the residual as 
\begin{equation*}
  r_{i} = \mathbf{x_{i}}^{T}\mathbf{\hat{\boldsymbol\beta}} - y_{i},
\end{equation*}
where $\hat{\boldsymbol\beta}$ is our estimate of $\boldsymbol\beta$. Then, we would like a Q-Q plot of the residuals with a Normal reference distribution to aid in testing assumption (4) above. Strictly speaking, even if assumption (4) is correct, the residuals are not independent because OLS leads to an estimation of $\hat{\boldsymbol\beta}$ such that $\sum_{i=1}^{n}r_{i} = 0$. However, accounting for the resulting negative correlation between the residuals when creating testing bands would be very difficult, and with a sufficiently large sample size this correlation becomes negligible. Thus, without an a priori reason to believe that the errors may deviate from normality at any particular point in the distribution, it makes sense to use ELL bands in this case. This is very easy to do with \pkg{qqconf}, as we show below.
\newline
\newline
First, we generate data to perform a regression. Here, we will generate each $\epsilon_{i}$ independently from a $t(2.5)$ distribution. 
%
<<generate_datas>>=
set.seed(100)
n <- 100
x <- runif(n)
eta <- rt(n, df = 2.5)
y <- x + eta
@
%
Then, we fit a regression with the simulated data
%
<<fit_regression>>=
reg <- lm(y ~ x)
@
%
With \pkg{base}-\proglang{R} functionality, we can can create a qqplot using the function \code{qqnorm} and \code{qqline}. 
%
%
<<base_qqplot, fig=TRUE>>=
qqnorm(reg$residuals)
qqline(reg$residuals)
@
%
%
\newline
Clearly there's some indication of deviation in this plot, but it can be hard to tell how significant the deviation is without confidence bounds. With our function \code{qq_conf_plot}, it is much easier to be able to judge the significance of the deviation.
%
%
<<qqconf_qqplot, fig=TRUE>>=
library(qqconf)
qq_conf_plot(
  obs = reg$residuals,
  samples_func = lines,
  samples_col = "red"
)
@
%
%
\newline
Here, we can clearly see that in the right tail of the residuals goes beyond the normal testing bounds, giving strong evidence that the errors were not generated from a normal distribution.

@@

\label{sec:discussion}

### 2021-06-08 article/manuscript.tex

A number of methods have been created towards this end. Naively, one could put pointwise testing bands onto Q-Q plots, which conduct $\alpha$ level tests on each order statistic of the sample. This method is clearly insufficient, as it ignores the multiple testing problem and thus provides no guarantee on the Type I error rate of testing the global null hypothesis that all of the data come from the specified null distribution. To appropriately deal with this multiple testing problem, the Komolgorov-Smirnov (KS) statistic is commonly used to create joint testing bands \citep{kolmogoroff1941confidence, smirnov1944approximate}. While this method ensures the correct Type I error by using the distribution of the KS statistic to create joint testing bands, this test suffers from very low power under a variety of reasonable alternatives since it has low sensitivity to deviation in the tails of the null distribution \citep{aldor2013power, berk1979goodness}. A more promising approach would be to apply the equal local levels method (ELL) to create testing bands. This method, originally introduced by Berk and Jones and later developed by Gontscharuk et al. as an alternative testing procedure to higher-criticism and KS, conducts a test with Type I error rate $\eta$ on each order statistic of the sample such that the Type I error rate of the global test is some desired level $\alpha$ \citep{berk1979goodness, gontscharuk2015intermediates, gontscharuk2016goodness, gontscharuk2017asymptotics}. Because this test conducts an $\eta$ level test on each order statistic, it can be used to create testing bands that are agnostic to any alternative distribution. While the ELL method has never previously been specifically applied to creating testing bands on Q-Q plots, we suggest it is a sensible choice for a standard test to use in creating testing bands on Q-Q plots, as it does not bias the user towards deviations of the sample in any particular part of the null distribution.

@@

Currently, there is no available software that allows a user to quickly create plausible and robust testing bands for Q-Q plots on real data. While the base-\proglang{R} package \pkg{stats} provides functionality for creating a Q-Q plot to compare a sample against the normal distribution, it doesn't give the user the ability to create Q-Q plots for other distributions or to put testing bands on those plots. The package \pkg{qqplotr} provides a number of helpful additions to the base-\proglang{R} functionality, including the ability to create Q-Q plots for a variety of reference distributions and a robust method for testing bands on the normal distribution \citep{qqplotr}. However, this package does not provide adequate testing bands for non-normal reference distributions and is not able to efficiently create testing bands for large sample sizes.

@@

%Add a line about Alder Noiman here where you mention tail sensitive and qqplotr
In what follows, we introduce the \proglang{R} package \pkg{qqconf}. With the aid of a novel recursive calculation developed below, \pkg{qqconf} can quickly provide ELL testing bands for comparing even very large samples to any reference distribution with a CDF implemented. In addition to these testing bands, \pkg{qqconf} provides a variety of plotting functionalities that allow the user to easily visualize where any deviation of the sample from the null distribution may occur. In section \hyperref[sec:methods]{2}, we introduce the methods required for the efficient recursive calculation mentioned above. In section \hyperref[sec:methods]{3}, we discuss the behavior of ELL bands in large sample sizes and an approximation to these bands that makes them very easy to calculate for even very large samples ($n > 5 \cdot 10^{6}$). In section \hyperref[sec:examples]{4}, we introduce \pkg{qqconf} and demonstrate its functionality on a few multiple testing scenarios including including Type I error calibration and genetic studies. Finally, we conclude in section \hyperref[sec:discussion]{5} with a discussion of limitations the of ELL and \pkg{qqconf} as well as further directions for creating testing bands on Q-Q plots.

@@

>>>>

A number of methods have been created towards this end. Naively, one could put pointwise testing bands onto Q-Q plots, which conduct an $\alpha$ level test on each order statistic of the sample. This method is clearly insufficient, as it ignores the multiple testing problem and thus provides no guarantee on the Type I error rate of testing the global null hypothesis that all of the data come from the specified null distribution. To appropriately deal with this multiple testing problem, the Komolgorov-Smirnov (KS) statistic is commonly used to create joint testing bands \citep{kolmogoroff1941confidence, smirnov1944approximate}. While this method ensures the correct Type I error by using the distribution of the KS statistic to create joint testing bands, this test suffers from very low power under a variety of reasonable alternatives since it has low sensitivity to deviation in the tails of the null distribution \citep{aldor2013power, berk1979goodness}. A more promising approach would be to apply the equal local levels method (ELL) to create testing bands. This global testing method, originally introduced by Berk and Jones and later developed by Gontscharuk et al. as an alternative testing procedure to higher-criticism and KS, conducts a test with Type I error rate $\eta$ on each order statistic of the sample such that the Type I error rate of the global test is some desired level $\alpha$ \citep{berk1979goodness, gontscharuk2015intermediates, gontscharuk2016goodness, gontscharuk2017asymptotics}. Previously, Alder-Noiman applied this method to create testing bands for Q-Q plots for the normal distribution using simulation \citep{aldor2013power}. Because this test conducts an $\eta$ level test on each order statistic and thus does not bias the user towards deviations of the sample from any particular part of the null distribution, we suggest it is a sensible choice for a standard method in creating testing bands for Q-Q plots.

@@

Currently, there is no available software that allows a user to quickly create plausible and robust testing bands for Q-Q plots on real data. While the base-\proglang{R} package \pkg{stats} provides functionality for creating a Q-Q plot to compare a sample against the normal distribution, it doesn't give the user the ability to create Q-Q plots for other distributions or to put testing bands on those plots. The package \pkg{qqplotr} provides a number of helpful additions to the base-\proglang{R} functionality, including the ability to create Q-Q plots for a variety of reference distributions and a robust method for testing bands on the normal distribution \citep{qqplotr}. Specifically, for the normal distribution \pkg{qqplotr} implements Alder-Noiman's simulation based method (which is referred to in the package as tail sensitive (TS) testing bands). However, this method does not provide testing bands for non-normal reference distributions and is not able to efficiently create testing bands for large sample sizes because it uses simulation.

@@

In what follows, we introduce the \proglang{R} package \pkg{qqconf}. With the aid of a novel recursive calculation developed below, \pkg{qqconf} can quickly provide ELL testing bands for comparing even very large samples to any reference distribution with a CDF implemented. In addition to these testing bands, \pkg{qqconf} provides a variety of plotting functionalities that allow the user to easily visualize where any deviation of the sample from the null distribution may occur. In section \hyperref[sec:methods]{2}, we introduce the methods required for the efficient recursive calculation mentioned above. In section \hyperref[sec:methods]{3}, we discuss the behavior of ELL bands in large sample sizes and an approximation to these bands that makes them very easy to calculate for even very large samples ($n > 5 \cdot 10^{6}$). In section \hyperref[sec:examples]{4}, we introduce \pkg{qqconf} and demonstrate its functionality on a few multiple testing scenarios including including Type I error calibration and genetic studies. Finally, we conclude in section \hyperref[sec:discussion]{5} with a discussion of some limitations of the ELL method and \pkg{qqconf} as well as further directions for creating testing bands on Q-Q plots.

@@

\subsection{Testing for Normality of Regression Residuals}
When performing an ordinary least squares regression (OLS), it is common to assume that the error terms are idependentally drawn from a normal distribution. Specifically, suppose that we have $n$ observations $\{\mathbf{x_{i}}, y_{i}\}_{i=1}^{n}$, where each $y_{i}$ is a scalar and each $\mathbf{x_{i}}$ is a $1xp$ vector of regressors. Then, we model each $y_{i}$ as
\begin{equation*}
  y_{i} = \mathbf{x_{i}}^{T}\boldsymbol\beta + \epsilon_{i}
\end{equation*}
where $\mathbf{\boldsymbol\beta}$ is a $1xp$ vector of regression parameters to be estimated and we assume
\begin{equation}
  \epsilon_{1}, ..., \epsilon_{n} \distras{iid} N(0, \sigma^{2}).
\end{equation}
Once $\boldsymbol\beta$ and $\sigma^{2}$ are estimated (typically via maximum liklihood), for each $y_{i}$ we define the residual as 
\begin{equation*}
  r_{i} = \mathbf{x_{i}}^{T}\mathbf{\hat{\boldsymbol\beta}} - y_{i},
\end{equation*}
where $\hat{\boldsymbol\beta}$ is our estimate of $\boldsymbol\beta$. Then, we would like a Q-Q plot of the residuals with a Normal reference distribution to aid in testing assumption (4) above. Strictly speaking, even if assumption (4) is correct, the residuals are not independent because OLS leads to an estimation of $\hat{\boldsymbol\beta}$ such that $\sum_{i=1}^{n}r_{i} = 0$. However, accounting for the resulting negative correlation between the residuals when creating testing bands would be very difficult, and with a sufficiently large sample size this correlation becomes negligible. Thus, without an a priori reason to believe that the errors may deviate from normality at any particular point in the distribution, it makes sense to use ELL bands in this case. This is very easy to do with \pkg{qqconf}, as we show below.
\newline
\newline
First, we generate data to perform a regression. Here, we will generate each $\epsilon_{i}$ independently from a $t(2.5)$ distribution. 
%
\begin{Schunk}
\begin{Sinput}
R> set.seed(100)
R> n <- 100
R> x <- runif(n)
R> eta <- rt(n, df = 2.5)
R> y <- x + eta
\end{Sinput}
\end{Schunk}
%
Then, we fit a regression with the simulated data
%
\begin{Schunk}
\begin{Sinput}
R> reg <- lm(y ~ x)
\end{Sinput}
\end{Schunk}
%
With \pkg{base}-\proglang{R} functionality, we can can create a qqplot using the function \code{qqnorm} and \code{qqline}. 
%
%
\begin{Schunk}
\begin{Sinput}
R> qqnorm(reg$residuals)
R> qqline(reg$residuals)
\end{Sinput}
\end{Schunk}
\includegraphics{manuscript-base_qqplot}
%
%
\newline
Clearly there's some indication of deviation in this plot, but it can be hard to tell how significant the deviation is without confidence bounds. With our function \code{qq_conf_plot}, it is much easier to be able to judge the significance of the deviation.
%
%
\begin{Schunk}
\begin{Sinput}
R> library(qqconf)
R> qq_conf_plot(
+    obs = reg$residuals,
+    samples_func = lines,
+    samples_col = "red"
+  )
\end{Sinput}
\end{Schunk}
\includegraphics{manuscript-qqconf_qqplot}
%
%
\newline
Here, we can clearly see that in the right tail of the residuals goes beyond the normal testing bounds, giving strong evidence that the errors were not generated from a normal distribution.
